Data irregularities across six implicit learning articles: Comments on Lola, Giatis, Pérez-Turpin, and Tzetzis (2021), Lola and Tzetzis (2021), Lola and Tzetzis (2020), Tzetzis and Lola (2015), Lola, Tzetzis, and Zetou (2012), and Tzetzis and Lola (2010)

### 2022-01-21 docs/manuscript/manuscript.Rmd

---
title             : "Data irregularities across six implicit learning articles: Comments on @lola2021, @lola2021b, @lola2020, @tzetzis2015, @lola2012, and @tzetzis2010"

@@

  \vspace{\baselineskip}\raggedright Submitted to *Meta-Psychology*. Participate in open peer review by commenting through hypothes.is directly on this preprint. The full editorial process of all articles under review at *Meta-Psychology* can be found here: https://tinyurl.com/mp-submissions You will find this preprint by searching for the first author's name.

authornote: |
  \vspace{-.5cm} 
  \addORCIDlink{Brad McKay}{0000-0002-7408-2323}

  \addORCIDlink{Michael J. Carter}{0000-0002-0675-4271}

abstract: |
  We present a critical re-analysis of six implicit learning papers published by the same authors between 2010 and 2021. We calculated effect sizes for each pairwise comparison reported in the papers using the data published in each article. We further identified mathematically impossible data reported in multiple papers, either with deductive logic or by conducting a GRIMMER analysis of reported means and standard deviations. We found the pairwise effect sizes were implausible in all six articles in question, with Cohen’s *d* values often exceeding 100 and sometimes exceeding 1000. \textcolor{red}{In contrast, the largest effect size observed in a million simulated experiments with a true effect of} *d* \textcolor{red}{= 3 was} *d*\textcolor{red}{ = 6.6.} Impossible statistics were reported in four out of the six articles. Reported test statistics and $\eta^{2}$ values were also implausible, with several $\eta^{2}$ = .99 and even $\eta^{2}$ = 1.0 for between-subjects main effects. The results reported in the six articles in question are unreliable. Many of the problems we identified could be spotted without further analysis, highlighting the need for adequate statistical training in the field of motor learning.

@@

Statistical reporting errors may commonly occur in psychology articles [@brown2017; @nuijten2016] and such errors are often consistent with hypothesized results [@bakker2011]. When the primary conclusions in research articles depend on reporting errors, replicability is unlikely and future research may be wasted if researchers attempt to build on the erroneously reported results [@munafo2017]. In this paper, we scrutinize six papers published by the same two authors[^1], where the authors report a high number of erroneous or implausible data on which their primary conclusions depend. We first became aware of the @lola2021 paper when the paper was highlighted in a social media post [@gray2021]. During an initial read through by one of us (BM), a number of reporting and/or statistical issues were noticed. The paper also referenced past research published by these authors. Given our cause for concern over the issues found in the @lola2021 paper, we deemed it necessary to examine these other papers. We remain agnostic to the sources of error in each of these papers. Nevertheless, the data irregularities we found are similar across the target articles and at times even include repeated values (e.g., $F$-statistics) across multiple papers. Regardless of the conclusion one reaches with respect to the mechanism behind these errors, it is our contention that the results reported in these papers are unreliable and that the respective journals in which the papers are published should take corrective actions[^2]. Below, we outline our causes for concern and the overarching issues we found across the six articles in question.

[^1]: One of the six papers had a third author and one had a third and fourth author.
[^2]: We contacted the Editors of the journals that published Articles 1, 3, 4, and 5 on Sept 22 2021. The Editors have either indicated their intention to further investigate these issues and/or take corrective actions.

@@

We reanalyzed six articles by Afroditi Lola, George Tzetzis, and their colleague [@lola2012; @lola2020; @lola2021; @lola2021b; @tzetzis2010; @tzetzis2015]. All six papers described experiments that sampled young females who were enrolled in a volleyball camp (see Table \@ref(tab:table1)). In their experiments, the authors investigated the effects of implicit and explicit instructions on learning motor and perceptual skills. Our reanalysis of the target articles evaluated the plausibility of the reported means, standard deviations, and test statistics. We will refer to the six articles throughout this paper using the following numbering system:

@@

[^3]: This article is in a journal of a publishing group that has been identified as a potential predatory journal. Further, we were unable to find an archived version of this article on the journal's webpage and were only able to find a posted version on ResearchGate (https://www.researchgate.net/profile/Angela-Calder/publication/234000504_The_scientific_basis_for_recovery_training_practices_in_sport/links/5428fff80cf26120b7b574ad/The-scientific-basis-for-recovery-training-practices-in-sport.pdf#page=57). In fact, the earliest available issue on the journal's webpage is from 2016.

\vspace{6ex}

@@

kbl(table_1,
    booktabs = TRUE,
    escape = FALSE,
    caption = "Participant demographics in each of the target articles.",
    col.names = c("Target article",
                  "Sample size and participant details"),
    align = c("l", "l")) %>%
  kable_styling(position = "left",
                font_size = 11) %>%
  pack_rows("Article 1", 1, 1) %>%
  pack_rows("Article 2", 2, 2) %>%
  pack_rows("Article 3", 3, 3) %>%
  pack_rows("Article 4", 4, 4) %>%
  pack_rows("Article 5", 5, 5) %>%
  pack_rows("Article 6", 6, 6) %>%
  footnote(alphabet = c("Articles 2, 3 and 4 report identical means and standard deviations for the age of their participants despite a different sample size in Article 3 from Articles 2 and 4, and a different age range in Article 4 from Articles 2 and 3."),
           alphabet_title = "Note.",
           threeparttable = TRUE)

@@

Although there were some differences between the reported experiments in the target articles, there were many methodological commonalities that can be summarized. All six articles involved female children learning a volleyball skill as part of a volleyball camp. In each case, the participants were reported to have minimal experience (i.e., were described as novices) with the task at hand. The purpose of all six experiments was to evaluate perceptual or motor learning differences as a function of the type of instruction received during practice. Each experiment included a pre-test, an acquisition (i.e., practice) phase involving 12 sessions spaced over four weeks, and a post-test. The authors also included a high stress test in Articles 1 [@lola2021b], 2 [@lola2021] and 6 [@tzetzis2010}.

In Articles 1-4, the groups differed with respect to the type of instruction received: implicit, explicit, or analogy. In Articles 5 and 6, a sequential group (see below for description) replaced the analogy group. All six experiments also included a control group that did not practice the task. Implicit instruction did not contain any explicit information for how to perform the task and the learners were asked to perform a distracting task like counting backwards while practicing to prevent them from acquiring declarative rules for performance. In contrast, explicit instruction consisted of direct verbal instructions for performing the task. Analogy instruction was considered a type of implicit instruction wherein an analogy or metaphor was provided to the learner. For example, "Imagine that the opponents' surface is covered with water. Send the ball where there is more water and no opponents at the court." [@lola2021, p. 9]. Sequential instruction involved receiving explicit instruction for the first half of training followed by implicit instruction for the second half of training. Across experiments, the authors predicted that implicit forms of instruction---implicit, analogy, and sequential---would be more effective than explicit instruction for motor and perceptual learning. This advantage was also predicted to be greater when testing was conducted in a high stress situation. In @lola2021 for instance, high stress was induced by falsely telling participants that the best performers would be selected for a draft to the national team. Further, it was predicted that analogy or sequential instruction would offer improvements relative to implicit instruction.

The primary outcome measures used in these experiments were reaction time (Articles 1, 2, 4, 5, and \textcolor{red}{6}), response accuracy (Articles 1, 2, 4, and 5), and motor performance measured on a 4-point scale (Article 3). In addition, Articles 2 and 6 included a measure of state anxiety, the Competitive State Anxiety Inventory-2 [@tsorbatzoudis1998], and Article 3 had a measure of self-efficacy using a Likert scale. Articles 2, 4, 5, and 6 also analyzed the number of explicit rules recalled.

@@

None of the six articles in question included a link to a public repository where the data could be accessed. We first wrote (email sent February 10, 2021) the corresponding author for Article 2 and asked if they would be willing to share the data for this experiment. The authors' response was that the data could not be shared as they were not finished with their analyses and were in the process of running different tests (A. Lola, personal communication, February 12 2021). We followed up this email (sent February 12 2021) by asking whether they would instead be willing to share the data from Articles 3-6 as these articles were less recent, and presumably all planned analyses had been completed. After a 2 week period with no response, we followed up with a third email (sent February 26 2021) and reiterated our interest in obtaining their data from Articles 3-6. The authors' response was that they were unable to share data from any of these articles because in some cases they no longer had the data and in other cases they had plans to conduct further analyses (A. Lola, personal communication, March 2 2021).

Our first two requests did not include any indication about our concerns regarding the data irregularities. Subsequently, in a fourth email (sent April 12 2021) we outlined our concerns for each article[^4] and once again reiterated our request to the authors to share any available data for any of the target articles. These requests were once again refused. The authors did address some specific concerns regarding Article 2, but for the most part only provided more general responses to our concerns. The authors admitted that some of the values reported in the other target articles were incorrect but did not identify which values or articles. Despite this, the authors maintained that the data irregularities---identified in our email and described in this paper---do not impact the veracity of their analyses or conclusions (A. Lola, personal communication, April 22 2021). We illustrate below that the data *and* analyses reported in each of the articles reviewed are unreliable. Our extracted data and analysis scripts can be accessed at the following link: https://osf.io/raz6q/.

@@

Means and standard deviations were extracted from each article for all measures and time points that were reported. Cohen's *d* was calculated for each pairwise comparison using the `compute.es` package in `R`. To provide context, we simulated data from two groups of *n* = 20, consistent with group sizes in Article 3, which had the largest groups among the target articles. We ran simulations with true effect sizes of *d* = .8 and *d* = 3 one million times each and report the range of effect sizes observed in those simulations.

@@

Each of the articles reported only omnibus test statistics and then reported post-hoc analyses with symbols demarcating significant and non-significant differences. In response to our expressions of concern, the authors suggested that many of the issues were due to misprints in the articles. Specifically, they indicated that the reported means and standard deviations in their tables were incorrect and the root of the errors had to be from them outsourcing the formatting of their tables. The authors then insisted that despite these typographic errors, their discussion of the results and corresponding conclusions were still accurate (A. Lola, personal communication, April 22, 2021). However, the test statistics reported for many analyses were implausibly large and the authors often reported $\eta^{2}$ values associated with the omnibus test. Our examination of the reported $\eta^{2}$ values revealed that, as with the reported pairwise comparisons, many were implausibly large.

@@

Cohen's *d* is used to describe the standardized mean difference of an effect and values can range between 0 and infinity \textcolor{red}{in both the negative and positive direction. We calculated absolute values so that all effects were positive}. Cohen's $d_{s}$ [@cohen1988] is the observed difference between group means divided by their pooled standard deviation [see @lakens2013 for a detailed discussion]. Conventional benchmarks for small, medium, and large effects are *d* = .2, .5, and .8, respectively [@cohen1962]; however, this *cargo-cult* approach to effect size interpretation has been heavily discouraged [see @correll2020; @field2016; @lakens2013; @thompson2007 for discussions]. Recently, an analysis of 6447 Cohen's *d* statistics extracted from social psychology meta-analyses observed median and 75th percentile Cohen's *d* values of .36 and .65, respectively---suggesting the conventional benchmarks may overestimate typical effects [@lovakov2021].\textcolor{red}{In the field of motor learning, recent meta-analyses have found average effect sizes in the published literature of} *d* = .19 [@mckay2021b], *d*= .54 [@mckay2021a], and *d* = .71 [@lohse2016].

To evaluate the maximum plausible Cohen's *d* statistics one might encounter from experiments similar to those reported in the target Articles, we conducted two simulations that each consisted of one million experiments (see Figure \@ref(fig:fig1)). We set the true effect size at *d* = .8, the conventional benchmark for a "large" treatment effect, in the first simulation. The largest effect size observed from the one million simulated experiments was *d* = 2.97. In the second simulation, we set the true effect size at *d* = 3, an unrealistically large effect size that might rarely be encountered in the psychology and/or motor learning literature. The maximum effect size observed in the one million simulated experiments was *d* = 6.6.

(ref:fig1-caption) Absolute Cohen's $d$ estimates from Articles 1-5 plotted on a logarithmic scale. Only data from the original tables in Article 2 are included. All pairwise comparisons have been included for all dependent measures in each experiment. The range of observed values from a simulation of 1,000,000 experiments with a true effect of $d$ = .8 is illustrated by shaded green and blue regions of the figure, reaching a maximum value of $d$ = 2.97. The range of observed values from a simulation of 1,000,000 experiments with a true effect of $d$ = 3 is illustrated by the shaded purple and blue regions of the figure, reaching a maximum value of $d$ = 6.6.

@@

In the context of the maximum values observed in our simulations, all six articles in question reported implausibly large effect sizes. \textcolor{red}{The original table of summary statistics in Article 1 indicated that} the smallest post-intervention difference in reaction times was *d* = .64. However, all other effects were larger than *d* = 8.7 and the largest effect was *d* = 41. The accuracy data also reflected improbably large post-intervention differences, with two-thirds of all comparisons showing effects larger than *d* = 5 and a largest effect of *d* = 13.35. \textcolor{red}{However, a correction to the tables of summary statistics was published that included substantially smaller standard deviations than the original tables. While the updated data do imply smaller effect sizes, as we discuss below, they appear to be inconsistent with the analyses reported in Article 2.}

In Article 2, the smallest pre-test difference for reaction time was *d* = 1.29 and the largest pre-test difference was *d* = 35.32---although none of the groups were reported as significantly different in the article. The smallest post-intervention effect at any of the three time points was *d* = 286.42, while the largest effect was *d* = 3504.86. A similar picture emerges when analyzing the accuracy data. All the pre-test differences were improbably large (all *d*'s $\geq$ 2.52) despite being reported as not significantly different in the articles. Ten of the pairwise comparisons resulted in *d*'s $\geq$ 100 following treatment with the independent variables. The motor component data revealed post-treatment effect sizes ranging from *d* = 1.16 to *d* = 13.5.

In Article 3, post-intervention motor performance effect sizes ranged from *d* = 3.1 to *d* = 20.95. Similarly, post-intervention self-efficacy effect sizes ranged from *d* = 1.79 to *d* = 44.46. Likewise, in Article 4 post-intervention reaction time effect sizes ranged from *d* = 2.28 to *d* = 35.97. Continuing this pattern, post-intervention response accuracy effect sizes ranged from *d* = 5.84 to *d* = 29.7.

In Article 5, many response accuracy effect sizes were implausibly large beginning at pre-test, wherein effects ranged from *d* = 2.53 to *d* = 15.50. Nevertheless, all pre-test comparisons were reported as non-significant. Following intervention, the effect sizes ranged from *d* = 23.13 to *d* = 155.08. Relative to other reported effect sizes, those reported for reaction time were not implausibly large at any time point, ranging from *d* = 0 to *d* = .86. However, the authors reported an implausibly large effect size, $\eta^{2}$ = .94, for the 4 (Group) x 3 (Time) ANOVA. Further, despite only one pairwise comparison being statistically significant, all post-intervention comparisons were reported as being significant in the article.

In Article 6, the authors did not report means and standard deviations for most of the analyses. However, $\eta^{2}$ effect sizes were reported and these ranged from $\eta^{2}$ = .52 to $\eta^{2}$ = .98. These effect sizes are discussed further below. All the post-intervention effects reviewed above were directionally consistent with the researchers' expectations. The sometimes implausibly large pre-test effects were not expected, but also were not reported as significant.

@@

In Article 2, the Competitive State Anxiety Inventory-2 was used to assess the level of cognitive and somatic stress experienced by participants. Responses were measured on a Likert scale ranging from 1 to 4 with the data appearing to represent the average response per item. At each of the three low-stress time points, the means reported for all four groups ranged from 1.02 to 1.09. During the high-stress time point, the means ranged from 3.95 to 4.09. The means for two groups were reported as greater than 4, which is not possible given the maximum score on the Competitive State Anxiety Inventory-2 is 4.

In Article 3, participants were asked to receive a served volleyball and pass it to a target consisting of three concentric circles. Motor performance was measured based on where the pass landed, with three points awarded for a pass to the central circle on the target, two points for the middle circle, one point for the outermost circle, and zero points for a pass that missed the target.[^5] Results were presented as average performance per trial and the analogy group was reported to have a mean score of 3.00 at retention (a perfect score) but with a standard deviation of .09. The perfect score was not a rounding error because the same group was reported to have a mean score of 2.99 with a standard deviation of .11 on the post-test. These data are not possible.

@@

Articles 2, 5, and 6 reported the means and standard deviations for the number of explicit rules recalled by participants following the intervention phase. As a single item analysis of integers these results were suitable for a GRIMMER analysis. In Article 2, the mean and standard deviation pairs were GRIMMER inconsistent for all four groups (I: *M* = .73, *SD* = .35; A: *M* = 1.03, *SD* = .25; E: *M* = 4.8, *SD* = .78; C: *M* = .67, *SD* = .48; *n* = 15). In Article 5, the mean and standard deviation pair was GRIMMER consistent for the explicit rules group (*M* = 4.8, *SD* = 1.78) \textcolor{red}{and the implicit group} (*M* = 2.3, *SD* = 1.3). The mean and standard deviation pairs for the remaining \textcolor{red}{two} groups were GRIMMER inconsistent (S: *M* = 4.2, *SD* = 1.07; C: *M* = 1.8, *SD* = .3; *n* = 15). In Article 6, the mean and standard deviation pairs were GRIMMER consistent for three of the four groups if the standard deviations were calculated for the population rather than the sample (S: *M* = 4.2, *SD* = 1.07; I: *M* = 2.3, *SD* = 1.3; C: *M* = 1.4, *SD* = .9; *n* = 12). For two of the groups, however, they were consistent regardless of which method of calculating the standard deviation was used. However, the results for the explicit group were GRIMMER inconsistent (*M* = 4.8, *SD* = 1.78).[^6]

[^6]: You may have noticed that \textcolor{red}{two} of the same mean and standard deviation pairings (*M* = 4.8, *SD* = 1.78; *M* = 4.2, *SD* = 1.07) were classified as GRIMMER inconsistent for one paper and consistent for the other. This is because of sample size differences (*n* = 15 and *n* = 12).

>>>>

---
title             : "Data irregularities across six implicit learning articles: Comments on @lola2021b, @lola2021, @lola2020, @tzetzis2015, @lola2012, and @tzetzis2010"

@@

  \vspace{\baselineskip}\raggedright Submitted to *Meta-Psychology*. Participate in open peer review by commenting through hypothes.is directly on this preprint. The full editorial process of all articles under review at *Meta-Psychology* can be found here: https://tinyurl.com/mp-submissions. You will find this preprint by searching for the first author's name.

authornote: |
  \vspace{-.5cm} 
  \addORCIDlink{Brad McKay}{0000-0002-7408-2323}
  \newline \indent \addORCIDlink{Michael J. Carter}{0000-0002-0675-4271}

abstract: |
  We present a critical re-analysis of six implicit learning papers published by the same authors between 2010 and 2021. We calculated effect sizes for each pairwise comparison reported in the papers using the data published in each article. We further identified mathematically impossible data reported in multiple papers, either with deductive logic or by conducting a GRIMMER analysis of reported means and standard deviations. We found the pairwise effect sizes were implausible in all six articles in question, with Cohen’s *d* values often exceeding 100 and sometimes exceeding 1000. \textcolor{red}{In contrast, the largest effect size observed in a million simulated experiments with a true effect of $d$ = 3 was $d$ = 6.6.} Impossible statistics were reported in four out of the six articles. Reported test statistics and $\eta^{2}$ values were also implausible, with several $\eta^{2}$ = .99 and even $\eta^{2}$ = 1.0 for between-subjects main effects. The results reported in the six articles in question are unreliable. Many of the problems we identified could be spotted without further analysis, highlighting the need for adequate statistical training in the field of motor learning.

@@

Statistical reporting errors may commonly occur in psychology articles [@brown2017; @nuijten2016] and such errors are often consistent with hypothesized results [@bakker2011]. When the primary conclusions in research articles depend on reporting errors, replicability is unlikely and future research may be wasted if researchers attempt to build on the erroneously reported results [@munafo2017]. In this paper, we scrutinize six papers published by the same two authors[^1], where the authors report a high number of erroneous or implausible data on which their primary conclusions depend. We first became aware of the @lola2021 paper when the paper was highlighted in a social media post [@gray2021]. During an initial read through by one of us (BM), a number of reporting and statistical issues were noticed. The paper also referenced past research published by these authors. Given our concerns over the issues found in the @lola2021 paper, we deemed it necessary to examine these other papers. We remain agnostic to the sources of error in each of these papers. Nevertheless, the data irregularities we found are similar across the target articles and at times even include repeated values (e.g., $F$-statistics) across multiple papers. Regardless of the conclusion one reaches with respect to the mechanism behind these errors, it is our contention that the results reported in these papers are unreliable and that the respective journals in which the papers are published should take corrective actions[^2]. Below, we outline our causes for concern and the overarching issues we found across the six articles in question.

[^1]: One of the six papers had a third author and one had a third and fourth author.
[^2]: On Sept 22 2021, we contacted the journal editors \textcolor{red}{of all articles except} @lola2021b \textcolor{red}{as we were unaware of this article at the time}. The Editors indicated their intention to further investigate these issues and/or take corrective actions. \textcolor{red}{On Jan 21 2022, we contacted the journal editor regarding the} @lola2021b \textcolor{red}{article.}

@@

\textcolor{red}{We reanalyzed the following six articles by Afroditi Lola and George Tzetzis:}

@@

[^3]: This journal has been identified as a potential predatory journal. We were unable to find an online version of this article on the journal's webpage and interestingly, the earliest issue on the webpage is from 2016. We were only able to find an online version on ResearchGate (https://www.researchgate.net/profile/Angela-Calder/publication/234000504_The_scientific_basis_for_recovery_training_practices_in_sport/links/5428fff80cf26120b7b574ad/The-scientific-basis-for-recovery-training-practices-in-sport.pdf#page=57).

\textcolor{red}{In all experiments, the authors investigated the effects of implicit and explicit instructions on perceptual and motor learning. All experiments sampled young females who were enrolled in a volleyball camp} (see Table \@ref(tab:table1)). \textcolor{red}{Our reanalysis of the target articles evaluated the plausibility of the reported means, standard deviations, and test statistics.}

\vspace{3ex}

@@

kbl(table_1,
    booktabs = TRUE,
    linesep = "",
    escape = FALSE,
    caption = "Participant demographics in each of the target articles.",
    col.names = c("Target article",
                  "Sample size and participant details"),
    align = c("l", "l")) %>%
  kable_styling(position = "left",
                font_size = 11) %>%
  footnote(alphabet = c("These articles report identical means and standard deviations for the age of their participants despite a different sample size in Lola & Tzetzis 2020 from Lola & Tzetzis 2021 and Tzetzis & Lola 2015, and a different age range in Tzetzis & Lola 2015 from Lola & Tzetzis 2020 and 2021"),
           alphabet_title = "Note.",
           threeparttable = TRUE)

@@

Although there were some differences between the reported experiments in the target articles, there were many methodological commonalities that can be summarized. All six articles involved female children learning a volleyball skill as part of a volleyball camp. In each case, the participants were reported to have minimal experience (i.e., were described as novices) with the task at hand. The purpose of all six experiments was to evaluate perceptual or motor learning differences as a function of the type of instruction received during practice. Each experiment included a pre-test, an acquisition (i.e., practice) phase involving 12 sessions spaced over four weeks, and a post-test. A high stress test was also included in @lola2021b, @lola2021, and @tzetzis2010.

In @lola2021b, @lola2021, @lola2020, and @tzetzis2015, the groups differed with respect to the type of instruction received: implicit, explicit, or analogy. In @lola2012 and @tzetzis2010, a sequential group (see below for description) replaced the analogy group. All six experiments also included a control group that did not practice the task. Implicit instruction did not contain any explicit information for how to perform the task and the learners were asked to perform a distracting task like counting backwards while practicing to prevent them from acquiring declarative rules for performance. In contrast, explicit instruction consisted of direct verbal instructions for performing the task. Analogy instruction was considered a type of implicit instruction wherein an analogy or metaphor was provided to the learner. For example, "Imagine that the opponents' surface is covered with water. Send the ball where there is more water and no opponents at the court." [@lola2021, p. 9]. Sequential instruction involved receiving explicit instruction for the first half of training followed by implicit instruction for the second half of training. Across experiments, the authors predicted that implicit forms of instruction---implicit, analogy, and sequential---would be more effective than explicit instruction for motor and perceptual learning. This advantage was also predicted to be greater when testing was conducted in a high stress situation. In @lola2021 for instance, high stress was induced by falsely telling participants that the best performers would be selected for a draft to the national team. Further, it was predicted that analogy or sequential instruction would offer improvements relative to implicit instruction.

The primary outcome measures used in these experiments were reaction time [@lola2021b; @lola2021; @tzetzis2015; @lola2012; @tzetzis2010], response accuracy [@lola2021b; @lola2021; @tzetzis2015; @lola2012], and motor performance measured on a 4-point scale [@lola2020]. In addition, @lola2021 and @tzetzis2010 included a measure of state anxiety, the Competitive State Anxiety Inventory-2 [@tsorbatzoudis1998], and @lola2020 had a measure of self-efficacy using a Likert scale. The number of explicit rules recalled was assessed in @lola2021, @tzetzis2015, @lola2012, and @tzetzis2010.

@@

None of the six articles in question included a link to a public repository where the data could be accessed. We first wrote (email sent February 10, 2021) the corresponding author of the @lola2021 paper and asked if they would be willing to share the data for this experiment. The authors' response was that the data could not be shared as they were not finished with their analyses and were in the process of running different tests (A. Lola, personal communication, February 12 2021). We followed up this email (sent February 12 2021) by asking whether they would instead be willing to share the data from their other articles [i.e., @lola2020; @tzetzis2015; @lola2012; @tzetzis2010] as these were less recent, and presumably all planned analyses had been completed. After a 2 week period with no response, we followed up with a third email (sent February 26 2021) and reiterated our interest in obtaining their data from any of these articles. The authors' response was that they were unable to share data from any of these articles because in some cases they no longer had the data and in other cases they had plans to conduct further analyses (A. Lola, personal communication, March 2 2021).

Our first two requests did not include any indication about our concerns regarding the data irregularities. Subsequently, in a fourth email (sent April 12 2021) we outlined our concerns for each article[^4] and once again reiterated our request to the authors to share any available data for any of the target articles. These requests were once again refused. The authors did address some specific concerns regarding @lola2021, but for the most part only provided more general responses to our concerns. The authors admitted that some of the values reported in the other target articles were incorrect, but did not identify which values or articles. Despite this, the authors maintained that the data irregularities---identified in our email and described in this paper---do not impact the veracity of their analyses or conclusions (A. Lola, personal communication, April 22 2021). We illustrate below that the data *and* analyses reported in each of the articles reviewed are unreliable. Our extracted data and analysis scripts can be accessed at the following link: https://osf.io/raz6q/.

@@

Means and standard deviations were extracted from each article for all measures and time points that were reported. Cohen's *d* was calculated for each pairwise comparison using the `compute.es` package in `R`. To provide context, we simulated data from two groups of *n* = 20, consistent with group sizes in @lola2020, which had the largest groups among the target articles. We ran simulations with true effect sizes of *d* = .8 and *d* = 3 one million times each and report the range of effect sizes observed in those simulations.

@@

Each of the articles reported only omnibus test statistics and then reported post-hoc analyses with symbols demarcating significant and non-significant differences. In response to our expression of concern, the authors suggested that many of the issues were due to misprints in the articles. Specifically, they indicated that the reported means and standard deviations in their tables were incorrect and the root of the errors had to be from them outsourcing the formatting of their tables. The authors then insisted that despite these typographic errors, their discussion of the results and corresponding conclusions were still accurate (A. Lola, personal communication, April 22, 2021). However, the test statistics reported for many analyses were implausibly large and the authors often reported $\eta^{2}$ values associated with the omnibus test. Our examination of the reported $\eta^{2}$ values revealed that, as with the reported pairwise comparisons, many were implausibly large.

@@

Cohen's *d* is used to describe the standardized mean difference of an effect and values can range between 0 and infinity \textcolor{red}{in both the negative and positive direction. We calculated absolute values so that all effects were positive}. Cohen's $d_{s}$ [@cohen1988] is the observed difference between group means divided by their pooled standard deviation [see @lakens2013 for a detailed discussion]. Conventional benchmarks for small, medium, and large effects are *d* = .2, .5, and .8, respectively [@cohen1962]; however, this *cargo-cult* approach to effect size interpretation has been heavily discouraged [see @correll2020; @field2016; @lakens2013; @thompson2007 for discussions]. Recently, an analysis of 6447 Cohen's *d* statistics extracted from social psychology meta-analyses observed median and 75th percentile Cohen's *d* values of .36 and .65, respectively---suggesting the conventional benchmarks may overestimate typical effects [@lovakov2021]. \textcolor{red}{In the field of motor learning, recent meta-analyses have found average effect sizes in the published literature of} *d* = .19 [@mckay2021b], *d* = .54 [@mckay2021a], and *d* = .71 [@lohse2016].

To evaluate the maximum plausible Cohen's *d* statistics one might encounter from experiments similar to those reported in the target articles, we conducted two simulations that each consisted of one million experiments (see Figure \@ref(fig:fig1)). We set the true effect size at *d* = .8, the conventional benchmark for a "large" treatment effect, in the first simulation. The largest effect size observed from the one million simulated experiments was *d* = 2.97. In the second simulation, we set the true effect size at *d* = 3, an unrealistically large effect size that might rarely be encountered in the psychology and/or motor learning literature. The maximum effect size observed in the one million simulated experiments was *d* = 6.6.

(ref:fig1-caption) Absolute Cohen's $d$ estimates from all articles except @tzetzis2010 plotted on a logarithmic scale. Only data from the original tables in @lola2021b are included. All pairwise comparisons have been included for all dependent measures in each experiment. The range of observed values from a simulation of 1,000,000 experiments with a true effect of $d$ = .8 is illustrated by shaded green and blue regions of the figure, reaching a maximum value of $d$ = 2.97. The range of observed values from a simulation of 1,000,000 experiments with a true effect of $d$ = 3 is illustrated by the shaded purple and blue regions of the figure, reaching a maximum value of $d$ = 6.6.

@@

In the context of the maximum values observed in our simulations, all six articles in question reported implausibly large effect sizes. \textcolor{red}{The original table of summary statistics in} @lola2021b \textcolor{red}{indicated that} the smallest post-intervention difference in reaction times was *d* = .64. However, all other effects were larger than *d* = 8.7 and the largest effect was *d* = 41. The accuracy data also reflected improbably large post-intervention differences, with two-thirds of all comparisons showing effects larger than *d* = 5 and a largest effect of *d* = 13.35. \textcolor{red}{However, a correction to the tables of summary statistics was published that included substantially smaller standard deviations than the original tables. While the updated data do imply smaller effect sizes, as we discuss below, they appear to be inconsistent with the reported analyses.}

In @lola2021, the smallest pre-test difference for reaction time was *d* = 1.29 and the largest pre-test difference was *d* = 35.32---although none of the groups were reported as significantly different in the article. The smallest post-intervention effect at any of the three time points was *d* = 286.42, while the largest effect was *d* = 3504.86. A similar picture emerges when analyzing the accuracy data. All the pre-test differences were improbably large (all *d*'s $\geq$ 2.52) despite being reported as not significantly different in the articles. Ten of the pairwise comparisons resulted in *d*'s $\geq$ 100 following treatment with the independent variables. The motor component data revealed post-treatment effect sizes ranging from *d* = 1.16 to *d* = 13.5.

In @lola2020, post-intervention motor performance effect sizes ranged from *d* = 3.1 to *d* = 20.95. Similarly, post-intervention self-efficacy effect sizes ranged from *d* = 1.79 to *d* = 44.46. Likewise, in @tzetzis2015 post-intervention reaction time effect sizes ranged from *d* = 2.28 to *d* = 35.97. Continuing this pattern, post-intervention response accuracy effect sizes ranged from *d* = 5.84 to *d* = 29.7.

In @lola2012, many response accuracy effect sizes were implausibly large beginning at pre-test, wherein effects ranged from *d* = 2.53 to *d* = 15.50. Nevertheless, all pre-test comparisons were reported as non-significant. Following intervention, the effect sizes ranged from *d* = 23.13 to *d* = 155.08. Relative to other reported effect sizes, those reported for reaction time were not implausibly large at any time point, ranging from *d* = 0 to *d* = .86. However, the authors reported an implausibly large effect size, $\eta^{2}$ = .94, for the 4 (Group) x 3 (Time) ANOVA. Further, despite only one pairwise comparison being statistically significant, all post-intervention comparisons were reported as being significant in the article.

In @tzetzis2010, the authors did not report means and standard deviations for most of the analyses. However, $\eta^{2}$ effect sizes were reported and these ranged from $\eta^{2}$ = .52 to $\eta^{2}$ = .98. These effect sizes are discussed further below. All the post-intervention effects reviewed above were directionally consistent with the researchers' expectations. The sometimes implausibly large pre-test effects were not expected, but also were not reported as significant.

@@

In @lola2021, the Competitive State Anxiety Inventory-2 was used to assess the level of cognitive and somatic stress experienced by participants. Responses were measured on a Likert scale ranging from 1 to 4 with the data appearing to represent the average response per item. At each of the three low-stress time points, the means reported for all four groups ranged from 1.02 to 1.09. During the high-stress time point, the means ranged from 3.95 to 4.09. The means for two groups were reported as greater than 4, which is not possible given the maximum score on the Competitive State Anxiety Inventory-2 is 4.

In @lola2020, participants were asked to receive a served volleyball and pass it to a target consisting of three concentric circles. Motor performance was measured based on where the pass landed, with three points awarded for a pass to the central circle on the target, two points for the middle circle, one point for the outermost circle, and zero points for a pass that missed the target.[^5] Results were presented as average performance per trial and the analogy group was reported to have a mean score of 3.00 at retention (a perfect score) but with a standard deviation of .09. The perfect score was not a rounding error because the same group was reported to have a mean score of 2.99 with a standard deviation of .11 on the post-test. These data are not possible.

@@

In @lola2021, @lola2012, and @tzetzis2010, the authors reported means and standard deviations for the number of explicit rules recalled by participants following the intervention phase. As a single item analysis of integers these results were suitable for a GRIMMER analysis. In @lola2021, the mean and standard deviation pairs were GRIMMER inconsistent for all four groups (Implicit: *M* = .73, *SD* = .35; Analogy: *M* = 1.03, *SD* = .25; Explicit: *M* = 4.8, *SD* = .78; Control: *M* = .67, *SD* = .48; *n* = 15). In @lola2012, the mean and standard deviation pair was GRIMMER consistent for the explicit rules group (*M* = 4.8, *SD* = 1.78) \textcolor{red}{and the implicit group} (*M* = 2.3, *SD* = 1.3). The mean and standard deviation pairs for the remaining \textcolor{red}{two} groups were GRIMMER inconsistent (Sequential: *M* = 4.2, *SD* = 1.07; Control: *M* = 1.8, *SD* = .3; *n* = 15). In @tzetzis2010, the mean and standard deviation pairs were GRIMMER consistent for three of the four groups if the standard deviations were calculated for the population rather than the sample (Sequential: *M* = 4.2, *SD* = 1.07; Implicit: *M* = 2.3, *SD* = 1.3; Control: *M* = 1.4, *SD* = .9; *n* = 12). For two of the groups, they were consistent regardless of which method of calculating the standard deviation was used. However, the results for the explicit group were GRIMMER inconsistent (*M* = 4.8, *SD* = 1.78).[^6]

[^6]: You may have noticed that \textcolor{red}{two} of the same mean and standard deviation pairings (*M* = 4.8, *SD* = 1.78 and *M* = 4.2, *SD* = 1.07) were classified as GRIMMER inconsistent for one paper and consistent for the other. This is because of sample size differences (*n* = 15 and *n* = 12).
